
\documentclass [PhD] {uclathes}

\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\newcommand\Tau{\mathcal{T}}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc} % usually not needed (loaded by default)
\usepackage[usenames,dvipsnames]{xcolor}
\definecolor{linkcolor}{RGB}{0,0,102}
\usepackage{hyperref} 
\usepackage{placeins}
\hypersetup{
	pdfstartview = {XYZ},
	colorlinks,
	allcolors = linkcolor,
	bookmarksopen,
	bookmarksnumbered
}
\usepackage{setspace}
\usepackage{silence}
\WarningFilter{caption}{Unknown} % captionpackage complains about thesis document class
\usepackage{caption}
\captionsetup{font={stretch=1}}
\usepackage{tabularx, colortbl}
\definecolor{mygray}{gray}{0.9}
\def \figscale {0.28}
\usepackage{multicol}
\usepackage{soul}
\usepackage{longtable,booktabs}

\newcommand{\scomment}[1]{{{\textcolor{blue}{{[Sriram: #1]}}}}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Usually things live in separate flies.
%
% \input {prelim}                           % preliminary page info

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                      %
%                          PRELIMINARY PAGES                           %
%                                                                      %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title          {Examining the impact \\
                of Neanderthal DNA \\
                on modern human biology}
\author         {Christopher Raymond Robles}
\department     {Human Genetics}
% Note:  degreeyear should be optional, but as of  5-Feb-96
% it seems required or you get a year of ``2''.   -johnh
\degreeyear     {2022}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chair          {Sriram Sankararaman}
\member         {Eleazar Eskin}
\member         {Paivi Pajukanta}
\member         {Bogdan Pasaniuc}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\dedication     {\textsl{For my family}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\acknowledgments {Graduate school at UCLA has been both a rewarding and trying experience, and I would like to thank all of the people who have helped me along my journey and seen all the ups and downs.
I would first like to thank my advisor Sriram, who has been the most supportive advisor I could ever ask for. As Sriram's first student, I have seen the lab grow from just him and I in the basement of Boyer hall, to a large group of students in the new Engineering 6 building. I vividly remember my first year working with Sriram, where he stayed late at night with me in the old lab, going through his CRF code line by line with me, patiently explaining everything. I started in his lab with a limited knowledge of statistics and machine learning, and his constant support, patience, and teachings have helped me grow tremendously as a mathematician, bioinformatician, scientist, and researcher. I have been so lucky to work with such a great scientist, mentor, and person overall. 

I would also like to the thank my committee: Eleazar Eskin, for always encouraging me and showing me that research can be fun and rewarding. His CGSI community has helped me build a large network, and allowed me to informally interact with and discuss reserach problems with many people who I consider to not only be my colleagues, but friends. Paivi Pajukanta, for always being there for me when I had any sort of problem, she was just a phone call away. Bogdan, for all the great discussion we had in journal clubs and at different events, as well as some cool soccer games.

I would also like to thank my lab mates, Alec, Brandon, Leah, Mike, Arun, Ariel, Ali, Ruthie, Elior, Liat, Brian, Rob Brown, Nadav and Johnson. All those great late nights shooting hoops around while we were going crazy with trying to figure out our codes.

I would like to thank my partner Sheila Rahimi, whose support during my toughest times means the world to me. 
I would also like to thank my friends at UCLA, Cat, Eric, Nate, Wenting, and Adriana. You guys made me stay sane through all my ups and downs, just being able to grab a beer a talk at the pub. 

I would like to thank my sister, Danielle Robles, she always had my back, its me and you against the world.

Last and not least, I would like to thank my parents Raymond and Lorraine Robles. They sacrificed so much for me, just to get me where I am today. Without their love and support I could never be where I was. I hope to one day pay you guys back for all the unconditional love, support, and cheering you have provided me with for all time. I love you both so much.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vitaitem   {2007--2011}
                {B.S. in Bioengineering \textit{cum laude honors}, University of California Santa Cruz, Santa Cruz, CA}
\vitaitem   {2010}
                {Undergraduate Research Assistant, Broad institute of Harvard and MIT, Cambridge, MA.}
\vitaitem   {2011--2013}
                {NIH Academy Research Fellow, National Human Genome Research Institute (NHGRI),Bethesda, MD} 
\vitaitem   {2013--2014}
                {MHIRT Research Associate, University of Otago, Dunedin, New Zealand} 
\vitaitem   {2014--2015}
                {Bioinformatics Analyst,  Mount Sinai School of Medicine, New York, NY}
\vitaitem   {2015--2022}
                {Graduate Student Fellow, University of California Los Angeles, Los Angeles, California}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

        \publication{Xinzhu Wei*, \textbf{Christopher R. Robles*, }, Ali Pazokitoroudi, Andrea Ganna, Alexander Gusev, Arun Durvasula, Steven Gazal, Po-Ru Loh, David Reich and Sriram Sankararaman.``The lingering effects of Neanderthal introgression on human complex traits". \textit{ In submission. *Equally contributed}, \\
        \url{https://www.biorxiv.org/content/10.1101/2022.06.07.495223v1.full.pdf}}      
        \publication {Charleston W. K. Chiang, Serghei Mangul, \textbf{Christopher R. Robles}, Sriram Sankararam``A comprehensive map of genetic variation in the world's largest ethnic group - Han Chinese". \textit{Molecular Biology and Evolution}, Volume 35, Issue 11, 1 November 2018, Pages 2736–2750, \url{https://doi.org/10.1093/molbev/msy170}}
        \publication{Kristina M. Garske, David Z. Pan, Zong Miao, Yash V. Bhagat, Caroline Comenho, \textbf{Christopher R. Robles}, Jihane N. Benhammou, Marcus Alvarez, Arthur Ko, Chun Jimmie Ye, Joseph R. Pisegna, Karen L. Mohlke, Janet S. Sinsheimer, Markku Laakso, Päivi Pajukanta.``Reverse gene-environment interaction approach to identify variants influencing body mass index in humans".\textit{Nat Metab 1}, 630–642 (2019),\\ \url{https://doi.org/10.1038/s42255-019-0071-6}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\abstract       {Recent genomic studies have revealed that all present-day non-African populations inherit $1 - 4\%$ of their genetic ancestry from a population related to the Neanderthals. Due to the high divergence of Neanderthals and modern humans, this introgression event introduced many novel mutations into  non-African populations. Although introgression between archaic and modern humans have previously been documented, the biological consequences of these admixture events on modern humans are not fully understood.  In this thesis, we examine these introgressed mutations to help elucidate the biological differences between Neanderthals and modern humans, as well as determine the selective forces that have acted on our genomes in the approximately 50,000 years since Neanderthal introgression occurred.

We first examine the impact Neanderthal introgression has on a single phenotype, Major Depressive Disorder, in more than 10,000 Han Chinese individuals who were diagnosed using structured clinical interviews. A recent examination of Neanderthal informative mutations (NIMs) among a large cohort of Europeans showed that these markers explained some proportions of the phenotypic risk of depression. Our results found no association between the NIMs and MDD, as well as a difference in the variance explained by the NIMs upon the phenotype vs Europeans. These results demonstrate the impact due to Neanderthal ancestry on one type of trait, MDD, appears to differ between Han Chinese and Europeans. 

We expanded upon a single phenotype by looking at the impact of Neanderthal introgression on wide range of phenotypes in white British individuals in the UK Biobank. We present a fine mapping
strategy for prioritizing causal NIMs that enables the identification of sets of NIMs that can credibly exert influence on specific phenotypes. We identify NIMs in a number of functionally important genes, including a premature stop codon in the FCGR2A gene, and a start codon loss in COQ10, pinpointing introgressed alleles in immune-related genes that
could have and continue to modulate human phenotype.

We then propose a pipeline that is able leverage information from Neanderthal introgressed alleles in order to identify Human-specific mutations that rose to near fixation in modern human populations, and examine how they
impact modern human phenotypes. We discover two regions of the
genome that are confidently fixed-derived and have associations with a number of phenotypes, and are in close proximity to genes that are related to immunity
suggesting that these regions may have importance in development of immune response in modern humans. 

Finally, we look to highlight possible mechanisms for how complex traits evolved in human history by examining the genetic contribution of Neanderthal
ancestry. We performed forward-in-time population genetic simulations to model the evolution of Neanderthal
and non-Neanderthal alleles according to a demographic model relating modern humans and
Neanderthals, to try an understand the evolutionary models that could explain observations of enriched or depleted contributions of Neanderthal ancestry to phenotypic variance seen in real data. We develop simulations of models of directional, stabilizing and disruptive selection, as well as propose two modified directional and stabilizing selection models, and find that most models lead to NIM heritability that is comparable or lower than non-introgressed SNPs, however our two modified models are able to present an increase in NIM heritability, suggesting possible mechanisms for how complex traits evolved in human history by examining the genetic contribution of Neanderthal
ancestry.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\begin {document}
\makeintropages

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Ordinarily each chapter (at least) is in a separate file.
%
\input {chapter1/chapter1}                         % Chapter 1 of dissertation
\input {chapter2/chapter2}                         % Chapter 2
\input {chapter3/chapter3}                         % etc.
\input {chapter4/chapter4}
\input {chapter5/chapter5}
%\input {chapter6/chapter6}
 \typeout{}
\bibliography{chapter1/chapter1,chapter2/chapter2,chapter3/chapter3,chapter4/chapter4,chapter5/chapter5}
\bibliographystyle{unsrt}

\end {document}

